State of the Practice for Lattice Boltzmann Method Software

### 2022-01-18 StateOfPractice/Papers/LatticeBoltz_SOP-arXiv/LatticeBoltz_SOP.tex
Updates to SOP for LBM, focus on comparing to research software in general

    We analyze the state of software development practice for Lattice Boltzmann
    solvers by quantitatively and qualitatively measuring and comparing 24
    software packages for 10 software qualities (installability, correctness/
    verifiability, reliability, robustness, usability, maintainability,
    reusability, understandability, visibility/transparency and
    reproducibility). Our reproducible analysis method employs a measurement
    template (containing 103 measures that are manually and automatically
    extracted from the project repository) and developer interviews (consisting
    of a set of 20 questions). From the measurement results, we ranked the
    software using the Analytic Hierarchy Process (AHP), a multi-criteria
    decision making method appropriate for cases with a mix of qualitative and
    quantitative factors. Our ranking was roughly consistent with ranking the
    repositories by the number of GitHub stars, although the number one project
    by star count (Sailfish) is ranked 9th by our methodology. Our top three
    packages were ESPResSo, Ludwig, and Palabos (all in the top 10 by star
    count). We interviewed 4 developers to gain insight into their current pain
    points. Identified challenges include lack of development time, lack of
    funding, and difficulty with ensuring correctness.  Based on our assessment,
    we make the following recommendations for current and future LBM projects:
    provide a detailed user manual and tutorials, explicitly state the limits of
    the software, use user-friendly programming languages, consider a peer
    review process, communicate development and contribution information, and
    use continuous integration and project management tools.  \wss{Change the
    abstract to discuss how LBM compares to common practices and best practices.
    Highlight some future practices.}

@@

The use of continuous integration tools and techniques alludes to a more refined
development process where faults are isolated and better recognized. Only three of the packages (ESPResSo, Ludwig, Musubi) mentioned applying the practice of continuous integration in their development process.

@@

All of the top four AHP ranked packages, ESPREsSo, Ludwig, Palabos, and OpenLB,
have each of the commonly found artifacts, except only three of them have a
requirements specification or theory notes. Palabos is the only one of these
four packages that does not have an artifact from this category.

@@

Many of the software packages that were assessed were developed by teams of two
or more people. Their work needed to be coordinated and managed.
Table~\ref{tbl_tools} shows the types of project management tools that were
explicitly noted in the artifacts, web-pages, or interviews with the developers.
As with development tools and dependencies, it is possible that other types of
project management tools are used, but they were not visible in the artifacts to
which we had access.  Collaboration tools are noted as being used when
developing the software projects. Most often email and video conferencing is
used. Project management software was not explicitly mentioned, but it is
possible that some of the projects use such software. Many of the projects are
located on GitHub, and its developers use the platform to help manage their
projects, especially bug related issues. Most of the projects appear to use
change tracking and version control tools. They often use GitHub or GitLab for
this. One package (SunlightLB) uses CVS. Document generation tools are mentioned
in the artifacts of 12 of the projects. The tools Sphinx and Doxygen are
explicitly used in this capacity. \wss{look at what the other papers say are
commonly used tools and bring that information into this section.}

@@

A lack of software development experience was noted by the developer of TCLB.
Many of the team members on their project are domain experts and there can be a
steep learning curve before these members contribute good quality source code.
This problem has been mitigated by re-writing the code to more easily enable
future contributions.  On a related point, improving the software engineering
education or experience of developers is also an idea that was brought up by
several developers.  \wss{The Wiese paper and others talk about the background
of developers. The knowledge for developing research software primarily comes
from peers and through self-study, not from formal training
\citep{HannayEtAl2009}.  The thinking of the developer on the role of formal education matches the trend \citet{PintoEtAl2018} observes, where in their replication of a previous study \citept{HannayEtAl2009}, they show a growing interest in formal training (From 13\% of respondants in 2009 to 22\% in 2018).}

@@

Difficulties with ensuring correctness were also noted by several developers.
They indicated that tests are run on all new source code additions, testing both
individual modules and the system to verify correctness. These tests compare the
package output to known correct output using test cases. The developer of TCLB
commented that the amount of testing data that is needed for some cases is a
problem as free testing services do not offer capabilities to store and process
such large amounts of data, and in-house testing solutions needed to be created
to address this limitation. The solution for this has been to limit the size of
the testing problems, and to run tests in small batches with few iterations.
\wss{The Weiss paper alludes this this as well.  I think the Kelly and Sanders
paper likely talks about testing problems and correctness.  ``while many
scientists believe that software testing is important, a smaller number believe
they have sufficient understanding about testing concepts''
\citep{HannayEtAl2009}.}

>>>>

    We analyze the state of software development practice for Lattice Boltzmann
    solvers by quantitatively and qualitatively measuring and comparing 24
    software packages for 10 software qualities (installability, correctness/
    verifiability, reliability, robustness, usability, maintainability,
    reusability, understandability, visibility/transparency and
    reproducibility). Our reproducible analysis method employs a measurement
    template (containing 103 measures that are manually and automatically
    extracted from the project repository) and developer interviews (consisting
    of a set of 20 questions). From the measurement results, we ranked the
    software using the Analytic Hierarchy Process (AHP). Our ranking was roughly
    consistent with ranking the repositories by the number of GitHub stars,
    although the number one project by star count (Sailfish) is ranked 9th by
    our methodology. Our top three packages were ESPResSo, Ludwig, and Palabos
    (all in the top 10 by star count). We interviewed 4 developers to gain
    insight into their current pain points. Identified challenges include lack
    of development time, lack of funding, and difficulty with ensuring
    correctness.  Based on our assessment, we make the following recommendations
    for current and future LBM projects: provide a detailed user manual and
    tutorials, explicitly state the limits of the software, use user-friendly
    programming languages, consider a peer review process, communicate
    development and contribution information, and use continuous integration and
    project management tools.  \wss{Change the abstract to discuss how LBM
    compares to common practices and best practices. Highlight some future
    practices.  \wss{Follow the general research software trend of increasing
    use of version control systems and project management on GitHub or
    similar.}}

@@

The use of continuous integration tools and techniques alludes to a more refined
development process where faults are isolated and better recognized. Only three
of the packages (ESPResSo, Ludwig, Musubi) mentioned applying the practice of
continuous integration in their development process. \wss{Forward reference that
suggests more tools use CI (Section 7 Comparison of Principles, Process and
Methodologies to Research Software in General).}

@@

All of the top four AHP ranked packages, ESPREsSo, Ludwig, Palabos, and OpenLB,
have each of the commonly found artifacts, except only three of them have a
requirements specification or theory notes. \wss{Change to clarify that it is
theory notes, not a requirements specification.  Remove requirements spec from
table and put it as rare.}  Palabos is the only one of these four packages that
does not have an artifact from this category.  \wss{** Requirements **}
\wss{Discuss requirements here. Extra information from Peter: I interpreted it
to mean what the software should do. As noted in my report, this was not common,
especially finding any formal document like an SRS. I do not recall any package
having an SRS. Some packages did note in their manuals or artifacts some of the
points that would be found in an SRS document. For example, DL\_MESO notes
functional (``The following is a list of the features that DLMESO currently
supports....'') and non-functional (should work in both ``serial and parallel
running; it can be run on standalone machines, clusters and supercomputers..'')
requirements in its documentation (more on this below). Yes, there was often
some sort of a theory resource. Often it was simply a link or reference to the
theory used in the package. There was significant discrepancy in how much theory
was documented or referenced. Meanwhile, requirements were rare. A few packages
did note some of this information within other artifacts but it was not
comprehensive, and it was not presented in a SRS document. DL\_MESO has some of
the information that belongs in an SRS in its user manual. It is not
comprehensive or presented as it would be in an SRS, but it was probably the
best example of requirements. It has a nice subsection on Functionality.
\citet{Nguyen-HoanEtAl2010} showed requirements are the least commonly produced
type of documentation for scientific software in general.  LBM matches this
trend.}

@@

Many of the software packages that were assessed were developed by teams of two
or more people. Their work needed to be coordinated and managed.
Table~\ref{tbl_tools} shows the types of project management tools that were
explicitly noted in the artifacts, web-pages, or interviews with the developers.
As with development tools and dependencies, it is possible that other types of
project management tools are used, but they were not visible in the artifacts to
which we had access.  Collaboration tools are noted as being used when
developing the software projects. Most often email and video conferencing is
used. Project management software was not explicitly mentioned, but it is
possible that some of the projects use such software. Many of the projects are
located on GitHub, and its developers use the platform to help manage their
projects, especially bug related issues. Most of the projects appear to use
change tracking and version control tools. \wss{**Version Control**} \wss{This
contrasts with work from a little over 10 years ago that shows version control
was only used in 50\% of research software projects \citep{Nguyen-HoanEtAl2010}.
At that time \citet{Nguyen-HoanEtAl2010} noted an increase from previous levels
of version control software use.  Anecdotally this confirms our observations.}
They often use GitHub or GitLab for this. One package (SunlightLB) uses CVS.
Document generation tools are mentioned in the artifacts of 12 of the projects.
The tools Sphinx and Doxygen are explicitly used in this capacity. \wss{look at
what the other papers say are commonly used tools and bring that information
into this section.}

@@

A lack of software development experience was noted by the developer of TCLB.
Many of the team members on their project are domain experts and there can be a
steep learning curve before these members contribute good quality source code.
This problem has been mitigated by re-writing the code to more easily enable
future contributions.  On a related point, improving the software engineering
education or experience of developers is also an idea that was brought up by
several developers.  \wss{The Wiese paper and others talk about the background
of developers. The knowledge for developing research software primarily comes
from peers and through self-study, not from formal training
\citep{HannayEtAl2009}.  The thinking of the developer on the role of formal education matches the trend \citet{PintoEtAl2018} observes, where in their replication of a previous study \citep{HannayEtAl2009}, they show a growing interest in formal training (From 13\% of respondants in 2009 to 22\% in 2018).}

@@

Difficulties with ensuring correctness were also noted by several developers.
They indicated that tests are run on all new source code additions, testing both
individual modules and the system to verify correctness. These tests compare the
package output to known correct output using test cases. \wss{oracle}  The
developer of TCLB commented that the amount of testing data that is needed for
some cases is a problem as free testing services do not offer capabilities to
store and process such large amounts of data, and in-house testing solutions
needed to be created to address this limitation. The solution for this has been
to limit the size of the testing problems, and to run tests in small batches
with few iterations. \wss{The Weiss paper alludes this this as well.  I think
the Kelly and Sanders paper likely talks about testing problems and correctness.
``while many scientists believe that software testing is important, a smaller
number believe they have sufficient understanding about testing concepts''
\citep{HannayEtAl2009}.}

\wss{** BEGIN Testing and Verification **}
\citet{Nguyen-HoanEtAl2010} says the following:
\begin{itemize}
	\item 38 of 60 respondents performed unit tests
	\item 23 of 60 respondents performed integration tests
	\item ``In our results, the four most common activities involve comparisons
	with known, trusted, or prior data. Verification and peer review were three
	of the four least commonly performed activities.'' \wss{testing is not the
	only way to verify software.  This addresses the oracle problem.}
\end{itemize}
\wss{** END Testing and Verification **}

``The questions of what constitutes an adequate testing oracle for scientific software and what constitutes adequate test data are left open. Some of the validation testing we researched was running up against the practical limits; they were using whatever oracle they could scrape together from industry data, benchmarks, and expert judgment.'' \citep{KellyAndSanders2008}

\wss{Seems that LBM not concerned with the oracle problem?}
